---
title: "PSTAT 10: Worksheet 10"
subtitle: "YOUR LAST NAME, YOUR FIRST NAME: YOUR NETID"
author: "Collaborated with: LAST NAME, FIRST NAME, NETID of any collaboraters"
date: "Make sure you show your work to your TA once complete."
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt
---

\newcommand{\Prob}{\mathbb{P}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(digits = 4)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


---

# Instructions for Completion

- We are back to having only computing exercises on the worksheets. As such, you must complete the computing exercises by filling in your code on this `.Rmd` file.

\

---

# Exercises

## Exercise 0: YAML

Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 


\

---

## Exercise 1: Waiting

The time it takes a department office to process a form is uniformly distributed between 4 days and 10 days. A form is selected at random, and the time it takes to get processed is recorded. \

**(a)** Define an appropriate random variable. What is the distribution of this random variable? _(Be sure to mention the parameters.)_

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answer.**


\

**(b)** What is the probability that it will take **exactly** 7 days to process the form?

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answer.**


\

**(c)** What is the probability that the office will take fewer than 8 days to process a randomly selected form?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

**(d)** What is the probability that the office will take between than 5 and 7 days to process a randomly selected form?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

---

## Exercise 2: Donuts!

Suppose the time time it takes a child to eat a donut is uniformly distributed between 0.5 and 4 minutes inclusive. \

**(a)** Calculate the probability that the child will take more than 3 minutes to eat the donut.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\


**(b)** Calculate the probability that the child will take between 0.75 and 1.5 minutes to eat the donut.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\


**(c)** Calculate the probability that a randomly selected child takes more than two minutes to eat a donut **given** that the child has already been eating the donut for more than 1.5 minutes.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\


---

## Exercise 3: Trees

You visit a national park and are informed that the height of a certain species of tree found in the forest is uniformly distributed between 3 and 70 feet. \


**(a)** What is the probability you encounter a tree shorter than 5.5 feet?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\


**(b)** For this probability density function, what is the height that marks the cutoff point of the tallest 15 percent of trees?

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\


**(c)** Evaluate the mean and standard deviation of the tree height distribution.

*Hint*: For a uniform distribution, the variance is $\sigma^2 = (b-a)^2/12$.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\



**(d)** Using (c), confirm that the chance that you encounter a tree with a height that is within half a standard deviation (that is, below or above) of the mean height is roughly 28.9 percent.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\


---

## Exercise 4: Writing Files

By now, we're experts at reading *in* files. One equally useful skill that we haven't really developed yet is that of *writing* files. As Data Scientists it is especially important that we are able to write files that we can give to our clients- clients would much prefer to receive an `.csv` file as opposed to a huge printout of numbers! In this problem, we'll get some practice with writing files.  \

**(a)** Look up the the help file for the function `write.csv()`, and comment on what you find.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```

**Replace this line with your answer.**


\

**(b)** Now, simulate 1 million draws from the $\mathrm{UNIF}(10.382 \ , \ 55.302)$ distribution; store these simulated draws in a vector called `draws_from_unif`.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

**(c)** Next, use the `write.csv()` function to export the values in your `draws_from_unif` variable to a file called `W10_unif_draws_data.csv`. 

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

**(d)** Finally, read in your `W10_unif_draws_data.csv` file, and use `head()` to display the first few elements in the file. Check that these match the first few values of the `draws_from_unif` variable you created in part (b) above.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

**As an Aside:** When you read your file back in, you might notice that `R` automatically exported the row names of your `draws_from_unif` vector. See if you can play around with the arguments to `write.csv()` to see if you can export your simulated draws *without* including the row names.

\

---

# Last exercise : Learning Gains

Please remember to fill out these learning gains before showing your TA your work to get Worksheet credit!

-   Gain 1
-   Gain 2
-   Gain 3



