---
title: "Your Turn : Named continuous distributions"
author: "Dr. Uma Ravat"
date: "`r Sys.Date()`"
tutorial:
  id: "Lecture09_Named distributions"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
description: "Named distributions"
runtime: shiny_prerendered
---


```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

##  Uniform  Distributions


A bus arrives at a stop every 10 minutes. A student is equally likely to arrive at the stop at any time. How long will the student have to wait?

What is the probability the waiting time, $X$, 

1. 5 minutes or less?
2. between 5 and 7 minutes?
3. more than 6 minutes?

###

- Let $X$ denote the waiting time until the next bus arrives.

- $X$ is a continuous uniform random variable, measured from 0 to 10 minutes.
- p.d.f is $f(x) = \frac{1}{10}, \text{if } 0 \leq x \leq 10$

It is always helpful(and mistakes are avoided) to **draw a picture** of the density and **shade the desired area** under the curve while doing probability calculations. 

###


### Probability calculation 1: Waiting time 5 minutes or less?



```{r waitingtime1, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r waitingtime1-solution}

punif(5, min = 0, max = 10)

```

### Probability calculation 2: Waiting time between 5 and 7 minutes?

```{r waitingtime2, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r waitingtime2-solution}

punif(7, min = 0, max = 10) - punif(5, min = 0, max = 10)

```

### Probability calculation 3: Waiting time more than 6 minutes?

```{r waitingtime3, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r waitingtime3-solution}
punif(10, min = 0, max = 10) - punif(6, min = 0, max = 10)
# or 
punif(6, min = 0, max = 10, lower.tail = FALSE)
# or
1 - punif(6, min = 0, max = 10)
```




## Generating uniform observations


### Generate a sample of size 5 from the uniform distribution on [0,10]

X ∼ Unif(0, 10)

Recall: 

- punif : cdf
- dunif : density
- runif : random sample
- qunif : quantile

```{r unif4, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r unif4-solution}
runif(5, min = 0, max = 10)
```

### Generate 5 samples and plot them 

```{r unif5, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r unif5-solution}
hist(runif(5, min = 0, max = 10))
```

### Generate many samples and plot them 

```{r unif6, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r unif6-solution}
hist(runif(100, min = 0, max = 10))
```



```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```

