---
title: "PSTAT 10: Worksheet 13"
subtitle: "YOUR LAST NAME, YOUR FIRST NAME: YOUR NETID"
author: "Collaborated with: LAST NAME, FIRST NAME, NETID of any collaboraters"
date: "Make sure you show your work to your TA once complete."
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

---

# Instructions for Completion

- We are back to having only computing exercises on the worksheets. As such, you must complete the computing exercises by filling in your code on this `.Rmd` file.

\

---

# Exercises

## Exercise 0: YAML

Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 


\

---

For this worksheet, we will again need to connect to the Chinook DB from lecture. Make sure the database is available to this .Rmd file in the `data` folder.

## Exercise 1: Establishing a Connection

Load the necessary packages and open a connection to the Chinook database housed in the sqlite file in the data folder.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

---

## Exercise 2: Terminology

You may use Google to find explanations that make sense to you. 

**(a)** Briefly explain the purpose of primary keys and foreign keys in a relational database.

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answers.**



\

**(b)** Briefly explain how relational databases model real world data.

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answers.**



\

**(c)** Briefly explain the difference between SQL, SQLite, and `RSQLite`.

`r bfcolor("Solutions:", "blue")`

**Replace this line with your answers.**


\

---


## Exercise 3: Tables and SQL queries

**(a)** List all tables in the database

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```

\

**(b)** Know the field names: What are the names of all the fields for `Customer`?


`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

**(c)** Look at the first five records for `CustomerId`, `FirstName`, `LastName` and `Address`.

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

**(d)** Select the `CustomerId`, `FirstName`, `LastName`, `State`, `Country` of all customers living in California.

_Hint_: Use `where State = 'CA'` in your select query

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

**(e)** How many customers are from USA? 

_Hint_: Use `count(*)` in your select query

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```


\

---

## Exercise 4: SQL queries

Choose another table from the database. Write a SQL query and describe in words what your query is doing. 

`r bfcolor("Solutions:", "blue")`

```{r}
# Write your code here

```

**Replace this line with your answers.**
\

---

# Last exercise : Learning Gains

Please remember to fill out these learning gains before showing your TA your work to get Worksheet credit!

-   Gain 1
-   Gain 2
-   Gain 3
