---
title: "L15-Rcode"
author: "Your netid, last name, first name here"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Connect to the database

```{r}
library(DBI)
library(RSQLite)
drv = dbDriver("SQLite") # the driver for the db you want to connect to
chinook_db = dbConnect(drv, # the driver to use
              dbname="./data/Chinook_Sqlite.sqlite") #path to the db file
```

## 2. Querying multiple tables : Joins

```{r}
dbGetQuery(chinook_db, "select TrackId, Name, Track.AlbumId, Title 
                          from Track inner join Album 
                             on Track.AlbumId = Album.AlbumId
                          where Track.AlbumId = 250
                          ")
```


## 3. YT with joins

In a single query, return the track id, track name, album id, mediatype id, and mediatype name of all tracks with AlbumId = 10.

```{r}
# Your code here
```



```{r}
dbGetQuery(chinook_db, 
"select Track.TrackId, Track.Name, AlbumId, MediaType.MediaTypeId, MediaType.Name
from Track inner join MediaType
    on Track.MediaTypeId = MediaType.MediaTypeId
where AlbumId = 10")
```

## 4. A Three-way Join
Return the track information and artist information 
of all tracks with AlbumId = 10.

```{r}
# Your code here
```


```{r}
dbGetQuery(chinook_db, 
"select TrackId, Track.Name, Track.AlbumId, Title, Artist.Name
 from Track
inner join Album on Track.AlbumId = Album.AlbumId
inner join Artist on Album.ArtistId = Artist.ArtistId
where Track.AlbumId = 250
                          ")
```

## 5. Table Alias

```{r}
dbGetQuery(chinook_db, 
              "select TrackId, t.Name, t.AlbumId, Title, ar.Name
               from Track t
              inner join Album al on t.AlbumId = al.AlbumId
              inner join Artist ar on al.ArtistId = ar.ArtistId
              where t.AlbumId = 250
                          ")
```


```{r}
dbGetQuery(chinook_db, 
"select TrackId, t.Name as TrackName, t.AlbumId,
Title as AlbumTitle, ar.Name as AristName
from Track t
inner join Album al on t.AlbumId = al.AlbumId
inner join Artist ar on al.ArtistId = ar.ArtistId
where t.AlbumId = 250
limit 5")
```

## 6. YT: A Three-way Join

```{r, echo = F}
# Produce the following relation with a three-way join.
#)

# Your code here

```



```{r, echo = T}
dbGetQuery(chinook_db,
"select TrackId, t.Name as TrackName, t.AlbumId,
Title as AlbumTitle, g.Name as GenreName
from Track t
    inner join Album al on t.AlbumId = al.AlbumId
    inner join Genre g on g.GenreId = t.GenreId
where t.AlbumId = 250
limit 5")
```

##  7. YT: Aggregation with Join

\tiny
```{r, echo = F}
#Retrieve the average track size  for all tracks in each genre,
#along with the name of the genre.

# Your code here

```



```{r}
dbGetQuery(chinook_db, 
"select avg(Bytes)  as AvgBytes,
       g.GenreId, g.Name as GenreName
from Track t
   inner join Genre g on g.GenreId = t.GenreId
  group by t.GenreId")
```


## 8. More on joins

`inner join` can be abbreviated with `join`. We don't need the word `inner`.

```{r}
dbGetQuery(chinook_db, 
"select TrackId, Name, t.AlbumId, Title 
from Track t
    join Album al on t.AlbumId = al.AlbumId
where t.AlbumId = 250
                          ")
```

##  9. *implicit* joins

The query below has no `join` keyword!
\tiny
```{r}
dbGetQuery(chinook_db, 
"select al.title, al.artistid, ar.artistid, ar.name
 from Album al, Artist ar
 where al.artistid = ar.artistid
limit 3")
```

##  10. `natural join`
**Avoid** natural joins.specify your keys explicitly.

```{r}
dbGetQuery(chinook_db, 
"select * from album natural join artist limit 3")
```



## 11. Disconnecting from the database

```{r eval=TRUE}
dbDisconnect(chinook_db)
```

```{r eval = F, error = T}
# Try selecting data
dbGetQuery(chinook_db, 
           "select CustomerId, FirstName, LastName from Customer")
```


