---
title: 'PSTAT 10: Homework 3'
subtitle: 'Dr. Ravat'
date: "Due on Canvas as per Canvas deadline"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F, message = F, warning = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}

library(kableExtra)
library(tidyverse)
```


---

# Instructions

- Remember to always copy the necessary files to your `workingfiles` folder; never edit and/or knit the files directly in the `10-content` folder.

- Rename your `.Rmd` file as `Homework03_your-netid.Rmd`. For example, my Rmd script file will be `Homework03_umaravat.Rmd`.

- Make sure to include your identifying details in the Rmd file.

- When submitting to Canvas, follow the provided submission instructions carefully.

`r bfcolor("Make sure you print relevant objects so that it is easy to spot whether you completed the exercise successfully or not. You won't get points if we can't verify this from your html file alone. Remove this paragraph, when you have satisfactorialy completed  all exercises and checked that the html file allows graders to verify your successful completion of exercises. Thank you! ", "red")`

---

# Exercises  


## Exercise 0: YAML


Update the YAML to reflect your own information (i.e. name, netid, etc.).



\

## Exercise 0.1: Installation of Requisite Package

`r bfcolor("Before knitting this document", "red")`, you will need to install the package `kableExtra`.
```{r, include=FALSE}
library(kableExtra)


```


\

---

## Exercise 1: Glossary of Functions

Fill in the following table with the functions you learned in Lecture 03 and 04 this week. Add as many lines as you feel are necessary (remember that Worksheet 0 covered how to add rows to tables in Markdown).

`r bfcolor("Solutions:", "blue")`

**NOTE:** Functions listed in the table may not be entirely comprehensive from the content covered in lecture 03 and 04.


| *`R` Function* | *Description of Function* | *Example of the Function* |
|------|-------|-------|
| `dim` | retrieve or set the dimension of an object | `dim(mtcars)` |
| `matrix` | creates a matrix (two-dimensional vector) from the given set of values | `matrix(1:9, 3, 3)` |
| `array` | creates an array, which can have one, two, or more dimensions | `array(1:16, dim = c(2,4,2))` |
| `dimnames` | retrieve or set the names (by dimension) of an object | `dimnames(A) <- list(c("r1","r2"),c("c1","c2"))` |
| `apply` | returns a vector or array or list of values obtained by applying a function to margins of an array or matrix | `apply(arr1, 2, sum)` |
| `factor` | used to encode a vector as a factor (essentially categories) | `factor(c("male","male",female",female"))` |
| `which` | give the TRUE indices of a logical object | `which(x > 0)` |
| `list` | creates a list, which allows for different data types and data structures to be included | `list(num = 1:9, log = TRUE)` |
| `data.frame` | creates data frames, where each column can be of a different data type | `data.frame(class = c("Math","Music"), score = c(85, 90))` |
| `any` | given a set of logical vectors, indicates if at least one of the values is true | `any(x > 0)` |
| `all` | given a set of logical vectors, indicates if all of the values are true | `all(x > 0)` |
| `nchar` | takes a character vector as an argument and returns a vector whose elements contain the sizes of the corresponding elements | `nchar("hello")` |
| `sub` | performs replacement of the first match in a string| `sub("l","w","hello")` |
| `gsub` | performs replacement of all matches in a string| `gsub("l","w","hello")` |










---


## Exercise 2: Thinking Critically About Data

As a Data Scientist, you may find yourself needing to code data into `R` manually (as opposed to simply reading in a file). In such cases, it is important to think critically about the data, and exactly *how* you want to encode it! \

As a simple example, we will consider the following dataset:

| Name | Major | Midterm Score (in \%) | Final Letter Grade |
|------|-------|---------------|--------------------|
| Brad | Statistics | 99 | A+ |
| Beckett | Statistics | 81 | A-
| Samanthan | Music | 94 | A+ |
| D'Vana | Biology |  75 | B- |
| Carol | Biology | 61 | C |

We will work toward inputting it as a data frame, but will do so by carefully considering each variable separately. \

\

**(a)** What Data Type should we use to encode the names of the individuals in the dataset? _No coding necessary for this part._


`r bfcolor("Solutions:", "blue")`

It is ideal to use the data type `character` (or `string`) to encode names. 


\

**(b)** Now, create a vector called `name` that stores the names of the individuals in the dataset.

`r bfcolor("Solutions:", "blue")`

```{r}
name <- c("Brad","Beckett","Samanthan","D'Vana","Carol")
name

```

Though this wasn't required, we can check the data type using `typeof()`:

```{r}
typeof(name)

```

\

**(c)** What Data Type should we use to encode the majors of the individuals in the dataset? _No coding necessary for this part._


`r bfcolor("Solutions:", "blue")`

Because there are several people with the same major, it is best to store majors in a `factor` data structure (which would correspond to an `integer` data type).


\

**(d)** Now, create a vector called `major` that stores the majors of the individuals in the dataset.

`r bfcolor("Solutions:", "blue")`

```{r}
major <- factor(c("Statistics","Statistics","Music","Biology","Biology"))
major

```
```{r}
## check data structure:
class(major)

## check data type:
typeof(major)


```


\

**(e)** What Data Type should we use to encode the midterm scores of the individuals in the dataset? _No coding necessary for this part._


`r bfcolor("Solutions:", "blue")`

Scores can be any numerical value, so we should use the `double` data type when storing exam scores. 



\

**(f)** Now, create a vector called `mt_scores` that stores the midterm scores of the individuals in the dataset.

`r bfcolor("Solutions:", "blue")`

```{r}
mt_scores <- c(99,81,94,75,61)
mt_scores
  
```
```{r}
## check the data type:
typeof(mt_scores)


```


\

**(g)** What Data Type should we use to encode the final letter grades of the individuals in the dataset? _No coding necessary for this part._


`r bfcolor("Solutions:", "blue")`

The data *structure* we should use is that of an ordered factor, which gives rise to an integer data *type*.

\

**(h)** Now, create a vector called `final_grades` that stores the final letter grades of the individuals in the dataset.

`r bfcolor("Solutions:", "blue")`

```{r}
final_grades <- factor(c("A+", "A-", "A+", "B-", "C" ),
                       ordered = TRUE,
                       levels = c("C", "B-", "A-", "A+"))

final_grades
```

```{r}
## check data structure:
class(final_grades)

## check data type:
typeof(final_grades)


```

\

**(i)** Now, it's time to put everything together! Using the variables you created in the above parts, encode the full dataset into a data frame called `df_dataset`

`r bfcolor("Solutions:", "blue")`

```{r}
df_dataset <- data.frame(name,major,mt_scores,final_grades)
df_dataset


```


\

**(j)** Suppose that Brad has pink hair; Beckett, Samanthan, and Carol have brown hair; and D'Vana has blue hair. Add a new column to `df_dataset` called `hair_color` containing this new information. Print `df_dataset` to ensure the column has successfully been added.

`r bfcolor("Solutions:", "blue")`

```{r}
df_dataset$hair_color <- factor(c("pink","brown","brown","blue","brown"))
df_dataset



```


\


---


# Exercise 3: Hairy Situ-eye-tion

Consider the data in the following table:

```{r, echo = F}
tab <- data.frame(
  C0 = c(rep("Hair Color", 3)),
  C1 = c("BLACK", "BROWN", "BLONDE"),
  C2 = c(32, 53, 03),
  C3 = c(11, 50, 30),
  C4 = c(10, 25, 05),
  C5 = c(03, 15, 08)
)

colnames(tab) <- c("", "", "BROWN", "BLUE", "HAZEL", "GREEN")

tab %>% 
  kbl(align = "c", caption = "") %>%
  add_header_above(c("Eye Color" = 6)) %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("bordered"), 
                full_width = F,
                position = "center",
                latex_options = "hold_position") %>%
  column_spec(1, border_left = T) %>%
  column_spec(6, border_right = T) %>%
  row_spec(0, bold = T, color = "blue") %>%
  column_spec(2, bold = T, color = "blue") %>%
  collapse_rows(columns = 1:2, valign = "middle")
```


**(a)** Construct a matrix using the data given in the table above, and assign it to the variable `hair_eye`. Fill in the matrix by rows.

`r bfcolor("Solutions:", "blue")`

```{r}
x <- c(32,11,10,3,53,50,25,15,3,30,5,8)
hair_eye <- matrix(x, nrow = 3, ncol = 4, byrow = TRUE)
hair_eye


```



\

**(b)** Add appropriate row names and column names 

`r bfcolor("Solutions:", "blue")`

```{r}
# Way 1 - using rownames and colnames
rownames(hair_eye) <- c("Black","Brown","Blonde")
colnames(hair_eye) <- c("Brown","Blue","Hazel","Green")
hair_eye

```

```{r}
# Way 2 - using dimnames()
dimnames(hair_eye) <- list(c("Black","Brown","Blonde"),
                           c("Brown","Blue","Hazel","Green"))
hair_eye

```


\

**(c)** Determine how many people have blonde hair and blue eyes

`r bfcolor("Solutions:", "blue")`

```{r}
# Way 1
hair_eye["Blonde","Blue"]

# Way 2
hair_eye[3,2]

```


\

**(d)** Sum row values, sum column values. _Hint:_ Check out the `rowSums` function in `R`

`r bfcolor("Solutions:", "blue")`

```{r}
rowSums(hair_eye)
colSums(hair_eye)


```

\

**(e)** Remove the column of data containing the hair colors of people with blue eyes. (After doing so, `hair_eye` should have one fewer columns)

`r bfcolor("Solutions:", "blue")`

```{r}
hair_eye <- hair_eye[,-2]
hair_eye


```


\


**(f)** It is found that there was a mistake when the data was collected: of the individuals with green eyes, 5 of them had black hair, 20 of them had brown hair, and 10 of then had blonde hair. Update your matrix `hair_eye` to reflect the corrected information.

`r bfcolor("Solutions:", "blue")`

```{r}
hair_eye[,3] <- c(5,20,10)
hair_eye



```


\


---


## Exercise 4: Homework *can* be Mean

In PSTAT 3029 (which doesn't exist... yet), the 10 homework scores for 3 students (named "Student 1", "Student 2", and "Student 3") are given as:

```{r}
hw_score <- list(s1 = runif(10, 50, 100),
                 s2 = runif(10, 50, 100),
                 s3 = runif(10, 50, 100))

```

**(a)** Retrieve the exam scores of Student 3 in two ways

`r bfcolor("Solutions:", "blue")`

```{r}
hw_score$s3

```

```{r}
hw_score[[3]]

```


\

**(b)** Describe the type of object returned by `hw_score[2]`

`r bfcolor("Solutions:", "blue")`

It is still a list!
```{r}
hw_score[2]

typeof(hw_score[2])
```


\

**(c)** Create a list `hw_score_average` that holds each students average hw score as its members. Name the averages as `ave_s1`, `ave_s2`, `ave_s3`. 

`r bfcolor("Solutions:", "blue")`

```{r}
hw_score_average <- list(avg_s1 = mean(hw_score$s1),
                         avg_s2 = mean(hw_score$s2),
                         avg_s3 = mean(hw_score$s3))
hw_score_average


```


\

**(d)** What is the overall average hw score? 

`r bfcolor("Solutions:", "blue")`

```{r}
mean(c(hw_score_average[[1]], hw_score_average[[2]], hw_score_average[[3]]))

```


\

---

## Exercise 5: Splitting Hairs... Again!

Recall that in homework 2 we examined the `strsplit()` function. In this problem, we investigate it a bit further. Additionally, for this problem it is important you perform the parts in order.

**(a)** Create a vector `x` and assign it the value `c("hair", "follicle")`. 

`r bfcolor("Solutions:", "blue")`

```{r}
x <- c("hair","follicle")
x

```


\


**(b)** What is the data type of `x`?

`r bfcolor("Solutions:", "blue")`

```{r}
typeof(x)

```


\


**(c)** Define `y <- strsplit(x, "i")`. What data structure does `y` possess?

`r bfcolor("Solutions:", "blue")`

```{r}
y <- strsplit(x, "i")

typeof(y)
class(y)

```


\

**(d)** What data type does `y[[2]]` possess?

`r bfcolor("Solutions:", "blue")`

```{r}
typeof(y[[2]])
class(y[[2]])

```


\

**(e)** Use subsetting on `y` to write a single line of code that returns the string `"foll"`. Don't just print `"foll"`!

`r bfcolor("Solutions:", "blue")`

```{r}
y[[2]][1]


```

\

---

## Exercise 6: More Palm-y Weather!

Recall the `palm.csv` dataset from Homework 2. 

\

**(a)** Copy the dataset from your HW02 `data` folder into a new folder (also called `data`) in your HW03 subdirectory. Add a statement to the fact that you completed the above exercise successfully or mention any difficulties you encountered or help received to complete this exercise.

`r bfcolor("Solutions:", "blue")`

I completed this exercise successfully.


\


**(b)** Load the data into your workspace, and store the resulting data in a variable called `palm_2`. 

`r bfcolor("Solutions:", "blue")`

```{r}
palm_2 <- read.csv("./data/palm.csv")

```


\

**(c)** Some of you noticed (on HW02) that there was no unique city block that contains the maximum number of Foxtail Palms. Run code to list **all** of the city blocks that contain the maximum number of Foxtail Palms represented in the dataset.

`r bfcolor("Solutions:", "blue")`

```{r}
which(palm_2$FP == max(palm_2$FP))

```


\

**(d)** Do any city blocks contain more than 8 King Palms? **Answer this question in TWO ways, using two DIFFERENT lines of code.**

`r bfcolor("Solutions:", "blue")`

```{r}
any(palm_2$KP > 8)

```

```{r}
which(palm_2$KP > 8)  # all these city blocks

```


\

## Learning Gains Week 02:

Write down your Learning Gains from section worksheets in Week 02

**(a)** Learning gains from the Worksheet 02. 

-   Gain 1
-   Gain 2
-   Gain 3

**(b)** Learning gains from the Worksheet 03. 

-   Gain 1
-   Gain 2
-   Gain 3



---