---
title: "Your Turn : Simulating Random variables in `R`"
author: "Dr. Uma Ravat"
date: "`r Sys.Date()`"
tutorial:
  id: "Lecture08_Random Variables"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
description: "Simulating Random variables in `R"
runtime: shiny_prerendered
---


```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Random variables, expected value, variance
### A card game 

In a game of cards you win \$1 if you draw a heart, \$5 if you draw an ace (including the ace of hearts), \$10 if you draw the king of spades and nothing for any other card you draw. 

### Exercise : Theory

Using R, tabluate the pmf of your winnings. Verify that it is a pmf. What is the expected value of your winnings? Interpret what this value means. 

```{r game-ev, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r game-ev-solution}
values <- c(0, 1, 5, 10)
p <- c(35/52, 12/52, 4/52, 1/52)
names(p) <- values
p
round(p,2)

sum(p)

ex <- sum(values*p)
ex
```

### Exercise: Simulation

Simulate playing this game 100 times. ie draw a sample of size 100 from the probability distribution of winnings from this game. View the results appropriately and find the average of the winnings in these 100 games. 

_Hint:_

- `sample()`: what values and with what probability, what's the sample size?
- What type of plot? (hint: Data falls into categories)


```{r game-sim100, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r game-sim100-solution}
values <- c(0,1,5,10)
p <- c(35/52, 12/52, 4/52, 1/52 )
names(p) <- values

sim_wins <- sample( values, 100, replace = TRUE, prob = p)
sim_wins

round(p,2)

table(sim_wins)

prop.table(table(sim_wins))

print(paste("average of winnings in 100 simulated games is", round(mean(sim_wins),2)))

print(paste("expect value of winnings  is", round(sum(values*p),2)))

barplot(table(sim_wins))
```

### Exercise

Compare the average winnings from simulation and theory. Are they same, different, how do you explain it?

B

### Exercise

What is the variability in winnings? What is its interpretation?

```{r game-var, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r game-var-solution}
values <- c(0, 1, 5, 10)
p <- c(35/52, 12/52, 4/52, 1/52 )
names(p) <- values

ev <- sum(values*p)
ev

v <-  sum(((values-ev)^2)*p)
v

sd <- sqrt(v)
sd

ev - sd

ev + sd

```

### Exercise

Find the cdf at 7. 

```{r game-cdf, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r game-cdf-solution}
values <- c(0,1,5,10)
p <- c(35/52, 12/52, 4/52, 1/52 )
names(p) <- values
p

cp <- cumsum(p)
cp

cp[3]
```

## Clinical Trial Simulation: pmf from simulation



Suppose that a clinical trial to test a new drug will be conducted on 8 patients, in which the probability of a good response to the drug is thought to be 0.15.

Let us simulate outcomes of 500 clinical trials, where each trial is conducted on 8 patients. 

The following code simulates the trial with 500 replicates.

```{r clinical-trial, exercise=TRUE, exercise.eval=FALSE}
#define parameters
number.patients = 8
response.prob = 0.15
number.replicates = 500

#create empty vectors to store results
number.responses.replicate = vector("numeric", number.patients)
number.responses = vector("numeric", number.replicates)

#set the seed for a pseudo-random sample
set.seed(10202022)

#simulate number.replicates many trials
for(k in 1:number.replicates){
  
  number.responses.replicate = sample(c(0,1), 
                                      size = number.patients,
                                      prob = c(1 - response.prob, response.prob),
                                      replace = TRUE)
  
  number.responses[k] = sum(number.responses.replicate)
  
}
print("Simulation ran successfully")
```

**(a)** Run the simulation and view the results (*Hint*: make a plot). Describe the distribution of the number of good responses to the drug. 


`r bfcolor("Solutions:", "blue")`
```{r clinical-trial1-setup}
#define parameters
number.patients = 8
response.prob = 0.15
number.replicates = 500

#create empty vectors to store results
number.responses.replicate = vector("numeric", number.patients)
number.responses = vector("numeric", number.replicates)

#set the seed for a pseudo-random sample
set.seed(10202022)

#simulate number.replicates many trials
for(k in 1:number.replicates){
  
  number.responses.replicate = sample(c(0,1), 
                                      size = number.patients,
                                      prob = c(1 - response.prob, response.prob),
                                      replace = TRUE)
  
  number.responses[k] = sum(number.responses.replicate)
  
}
```

```{r clinical-trial1, exercise=TRUE, exercise.eval=FALSE}

```
```{r clinical-trial1-solution}
table(number.responses)
barplot(table(number.responses))
```

**(b)** What value(s) for response probability would produce a left-skewed distribution? What value(s) would produce a symmetric distribution? (See code for a matrix of plots)

`r bfcolor("Solutions:", "blue")`

```{r clinical-trial2, exercise=TRUE, exercise.eval=FALSE}

```

```{r clinical-trial2-solution}
#define parameters
number.patients = 8
response.prob = 0.85
number.replicates = 500

#create empty vectors to store results
number.responses.replicate = vector("numeric", number.patients)
number.responses.A = vector("numeric", number.replicates)

#set the seed for a pseudo-random sample
set.seed(10202022)

#simulate the trials
for(k in 1:number.replicates){
  
  number.responses.replicate = sample(c(0,1), size = number.patients,
                                      prob = c(1 - response.prob, response.prob),
                                      replace = TRUE)
  
  number.responses.A[k] = sum(number.responses.replicate)
  
}

#define parameters
number.patients = 8
response.prob = 0.5
number.replicates = 500

#create empty vectors to store results
number.responses.replicate = vector("numeric", number.patients)
number.responses.B = vector("numeric", number.replicates)

#set the seed for a pseudo-random sample
set.seed(10202022)

#simulate the trials
for(k in 1:number.replicates){
  
  number.responses.replicate = sample(c(0,1), size = number.patients,
                                      prob = c(1 - response.prob, response.prob),
                                      replace = TRUE)
  
  number.responses.B[k] = sum(number.responses.replicate)
  
}

par(mfrow = c(1, 2))
barplot(table(number.responses.A))
barplot(table(number.responses.B))
```    

**(c)** Based on the results of the simulation, estimate the probability that 0 patients respond well to the new drug. 

`r bfcolor("Solutions:", "blue")`  

The probability that 0 patients respond well is 132/500 = 0.264

```{r clinical-trial3-setup}
#define parameters
number.patients = 8
response.prob = 0.15
number.replicates = 500

#create empty vectors to store results
number.responses.replicate = vector("numeric", number.patients)
number.responses = vector("numeric", number.replicates)

#set the seed for a pseudo-random sample
set.seed(10202022)

#simulate number.replicates many trials
for(k in 1:number.replicates){
  
  number.responses.replicate = sample(c(0,1), 
                                      size = number.patients,
                                      prob = c(1 - response.prob, response.prob),
                                      replace = TRUE)
  
  number.responses[k] = sum(number.responses.replicate)
  
}
```




```{r clinical-trial3, exercise=TRUE, exercise.eval=FALSE}

```
```{r clinical-trial3-solution}
sum(number.responses == 0)
sum(number.responses == 0)/number.replicates
```

**(d)** Let $X$ be a random variable defined as the number of patients who respond well to the experimental drug. Based on the results of the simulation,  construct a probability distribution for $X$, calculate $E(X)$, $V(X)$ and the c.d.f
    
`r bfcolor("Solutions:", "blue")`

$X$ can take on values 0 through 8, inclusive.

```{r clinical-trial4-setup}
#define parameters
number.patients = 8
response.prob = 0.15
number.replicates = 500

#create empty vectors to store results
number.responses.replicate = vector("numeric", number.patients)
number.responses = vector("numeric", number.replicates)

#set the seed for a pseudo-random sample
set.seed(10202022)

#simulate number.replicates many trials
for(k in 1:number.replicates){
  
  number.responses.replicate = sample(c(0,1), 
                                      size = number.patients,
                                      prob = c(1 - response.prob, response.prob),
                                      replace = TRUE)
  
  number.responses[k] = sum(number.responses.replicate)
  
}
```


```{r clinical-trial4, exercise=TRUE, exercise.eval=FALSE}

```

```{r clinical-trial4-solution}
#using r as a calculator
prop.table(table(number.responses))  #print probability distribution table
    
#calculate E(X)
xi = 0:8
prob.xi = prop.table(table(number.responses))
ex = sum(xi[1:6]*prob.xi) # try 1:8 and see what happens. Why?
ex

```


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```



