---
title: "Your Turn : Named Discrete distributions"
author: "Dr. Uma Ravat"
date: "`r Sys.Date()`"
tutorial:
  id: "Lecture09_Named distributions"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
description: "Named distributions"
runtime: shiny_prerendered
---


```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```


```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

##  Binomial Distributions

### Probability calculation 1. 

Let X ∼ Binom(10, 1/4). Determine the probability of getting 4 successes, i.e P(X = 4).

*Hint*: What parameters are we working with? What is our sample size? What is our probability of success? 

```{r bin1, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r bin1-solution}
# n <- 10
# p <- 1/4
# k <- 4
dbinom(4, size = 10, prob = 1/4)
```

### Plotting the pmf

The plot of the pmf of X is

```{r bin2, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r bin2-solution}
plot(0:10, dbinom(0:10, 10, 1/4), 
     xlab="Number of successes", ylab="Probability", type="h")
```

### Probability calculation 2

What’s the probability of at most 4 successes, i.e P(X ≤ 4) =?

```{r bin3, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r bin3-solution}
# method 1
sum(dbinom(0:4, size = 10, prob = 1/4))

# method 2
pbinom(4, size = 10, prob = 1/4)
```

### Plotting the cdf

The plot of the cdf of X is

```{r bin4, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r bin4-solution}
plot(0:10, pbinom(0:10, 10, 1/4), 
     xlab="Number of successes", ylab="Probability", type="s", lwd=4)
abline(0.53, 0, lty = 2, col = "red")
abline(0.78, 0, lty = 2, col = "red")
```

### Probability calculaton 3

Let X ∼ Binom(10, 1/4). What’s the probability of getting between 4 and 8 successes, i.e P(4 ≤ X ≤ 8) = ?

```{r bin5, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r bin5-solution}
# method 1
pbinom(8, 10, 1/4) - pbinom(3, 10 , 1/4)

# method 2
sum(dbinom(4:8, 10,1/4))
```

### Probability calculation 4

What’s the probability of getting more than 4 successes, i.e P(X > 4) = ?

```{r bin6, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r bin6-solution}
1 - pbinom(4, 10, 1/4)
```

## Generating a sample of binomial observations

If X ∼ Binom(10, 0.8) then X is total number of successes in 10 trials of a binomial experiment with p = 0.8

We can model this using `sum(sample(0:1, 10, replace = T, prob = c(0.2, 0.8)))` = `r sum(sample(0:1, 10, replace = T, prob = c(0.2, 0.8)))`

However, we can also use R's built in binomial generator:

```{r binE1, echo = T, exercise = FALSE}
rbinom(1, size = 10, prob = 0.8)
```

### Generate a sample of size 5 from the binomial distribution

X ∼ Binom(10, 0.8)

```{r bin7, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r bin7-solution}
rbinom(5, size = 10, prob = 0.8)
```

## Looking at the family of all binomial distributions

Binomial distributions for n = 10 and different p

```{r binE2, echo = T, exercise = FALSE}
par(mfrow = c(2,2))
plot(0:10,dbinom(0:10, 10, 0.1),
type="h", xlab = "x", ylab = "P(X=x)",
main = "p = 0.1")
plot(0:10,dbinom(0:10, 10, 0.3),
type="h", xlab = "x", ylab = "P(X=x)",
main = "p = 0.3")
plot(0:10,dbinom(0:10, 10, 0.7),
type="h",xlab = "x", ylab = "P(X=x)",
main = "p = 0.7")
plot(0:10,dbinom(0:10, 10, 0.9),
type="h", xlab = "x", ylab = "P(X=x)" ,
main = "p = 0.9")
```

### Generate binomial distributions for different values of n and same/different values of p 


```{r bin8, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```

```{r bin8-solution}
par(mfrow = c(2,2))
plot(0:10,dbinom(0:10, 10, 0.7),
type="h", xlab = "x", ylab = "P(X=x)",
main = "n = 10")
plot(0:100,dbinom(0:100, 50, 0.7),
type="h", xlab = "x", ylab = "P(X=x)",
main = "n = 100")
plot(0:1000,dbinom(0:1000, 100, 0.7),
type="h",xlab = "x", ylab = "P(X=x)",
main = "n = 1000")
```
What can you conclude from the two above exercises? This is what a data scientist sort of does - looks at data and draws some conclusions. 

### Try something else if you feel like

```{r try, exercise=TRUE, exercise.eval=FALSE}
your_code_here
```



```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```

